### 2007-01-11 Documentation/part2.tex
*** empty log message ***

\section{More advanced AEDL features}

@@

\index{aedl}{{\tt nuflux}!time@{\tt "@time}}
There are two principal ways to initialize a neutrino flux: 
Either one can use 
a built-in source, or one can provide a file. In both cases,
a flux is defined by the environment {\tt nuflux}, such as
\index{aedl}{{\tt nuflux}}
\begin{quote}
  {\tt nuflux(\#name)<\\
\tb $\ldots$\\
\tb @time = 8.0 \\
>}
\end{quote}
with a running time of $8$ years. Note that the running time is used within
the {\tt nuflux} environment. This feature can be used to load the neutrino and antineutrino fluxes separately, in order to combine them with different
running times within one experiment. The name of the flux {\tt \#name} will later be referred to in the channel definitions. Note that \GLOBES\ versions older than 3.0 use the still supported {\tt flux} environment, which is different from {\tt nuflux} by an undocumented normalization factor 5.2 for user-defined fluxes. This difference is explained in the appendix on page~\pageref{app:flux}.

\index{aedl}{{\tt flux}}

@@

For a built-in neutrino source, one has to specify which
built-in spectrum has to be used, as well as its parameters. The software
will then automatically calculate the neutrino spectrum. Note that in this
case, there is no degree of freedom in the choice of the source units.
The currently available fluxes are described in \Tab~\ref{tab:binfluxes}.
For example, two built-in neutrino factory fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}). In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year. Thus, an example to set up a neutrino factory flux is 
\index{aedl}{{\tt nuflux}!builtin@{\tt "@builtin}}
\index{aedl}{{\tt nuflux}!parent energy@{\tt "@parent\_energy}}
\index{aedl}{{\tt nuflux}!stored muons@{\tt "@stored\_muons}}
\index{aedl}{{\tt nuflux}!end point@{\tt "@end\_point}}
\index{aedl}{{\tt nuflux}!stored ions@{\tt "@stored\_ions}}
\index{aedl}{{\tt nuflux}!gamma@{\tt "@gamma}}
\begin{quote}
{\tt nuflux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}
\end{quote}
%
\index{aedl}{{\tt nuflux}!power@{\tt "@power}}
\index{aedl}{{\tt nuflux}!norm@{\tt "@norm}}
\index{aedl}{{\tt nuflux}!flux file@{\tt "@flux\_file}}
For a user-defined flux, one has to give it the file name:
\begin{quote}
{\tt nuflux(\#user)<}\\
{\tt \tb @flux\_file = "user\_file\_1.dat"\\
\tb @time = 2.0\\
\tb @power = 4.0\\
\tb @norm = 1e+8}\\
{\tt >}
\end{quote}
In this case, the {\tt @norm} variable is an overall normalization which defines a conversion factor from the fluxes in the file to the units in \GLOBES . In general, there are many ways to give the source power of a 
neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
not always straightfoward to calculate. Often, one has to take into account
many things, such as the number of target particles per unit mass. 
In addition, the fluxes will be rescaled by $1/L^2$, which means that the
normalization must contain a factor $L_0^2$. Here $L_0$ is the distance from the source for which the flux is given to the actual neutrino production region. At the end, it is left to the user to ensure that the 
numbers in the flux file give, after the multiplication with {\tt @norm}, 
the proper numbers of produced neutrinos corresponding to the chosen target power {\tt @power}. Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.
For more details on the flux definition, see appendix on page~\pageref{app:flux}.

\index{aedl}{{\tt flux}!flux file@{\tt "@flux\_file}}\index{norm}{Flux!file}
The software assumes that the given flux file has seven columns and
501 lines with equidistant energies. The format is:
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
In order to access fluxes at arbitrary energies, linear interpolation 
is used by the software. In general, it is advisable to provide the flux between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}), since these values are used by the software. However, if the energy leaves the range of values given in the file, zero is returned.
The the columns for the fluxes for unused flavours have to be filled all the 
same,\eg\ with zeros.

@@

Cross sections\index{norm}{Cross section} will later be used as part of the 
channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
fluxes, they are provided by the user as a data file:
\index{aedl}{{\tt cross}}
\index{aedl}{{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quote}
{\tt cross(\#name)<}\\
{\tt \tb @cross\_file ="user\_file\_1.dat"}\\
{\tt >}
\end{quote}  
This cross section can later be refered to by {\tt \#name}.

Cross sections in \GLOBES\ are given as total cross section divided by energy:
\begin{equation}
\hat\sigma(E)=\sigma(E)/E\,\left[ 10^{-38}\,
\frac{\mathrm{cm}^2}{\mathrm{GeV}} \right]
\end{equation}
The software assumes that the cross section files are text files with 
seven columns and $1001$ lines of the form
\index{norm}{Cross section!file}
\index{aedl}{{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quotation}
$\mathrm{log}_{10} E\quad
\hat\sigma_{\nu_e}\quad
\hat\sigma_{\nu_\mu}\quad
\hat\sigma_{\nu_\tau}\quad
\hat\sigma_{\bar\nu_e}\quad
\hat\sigma_{\bar\nu_\mu}\quad
\hat\sigma_{\bar\nu_\tau}$
\end{quotation}
Here the logarithms of the energy values have to be equidistant. For 
arbitrary energies, linear interpolation is used. If the energy leaves the
range of values given in the file, $0.0$ will be assumed. In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
Unused cross sections have to be filled with zeros, and can not
 be just omitted.

@@

\index{aedl}{{\tt channel}|(}
Channels\index{norm}{Channel} in \GLOBES\ represent an intermediate level 
between the pure oscillation physics given by the oscillation probability
$P_{\alpha\beta}$, and the total event rates composed of signal and 
background. A channel describes the path from one
initial neutrino flavor in the source to the event rates in the detector for one specific interaction type (IT) and final flavor.  
Therefore, a channel contains the description of the 
initial neutrino flavor, its CP eigenvalue 
(neutrino or antineutrino)\footnote{Currently \GLOBES\ does not support lepton
number violating transitions, \ie\ no transitions from neutrino to 
antineutrino (or vice versa) are considered.}, 
the detected neutrino flavor, the interaction cross sections for the chosen interaction type, and the energy resolution function of the detector.

@@

A simple channel definition in \GLOBES\ consists of the flux,
the CP-sign of the initial state, the initial flavor, the final flavor,
the cross sections, and the energy resolution function. In order to refer to
flux, cross sections, and energy resolution functions, they have to be 
defined before with their {\tt \#name} in the respective environments. 
Thus, a simple definition of a channel\index{aedl}{{\tt channel}} is
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}
%
\index{norm}{Oscillation!switching off}
It is also possible to define a channel as no-oscillation by using the
prefix {\tt NOSC\_}\index{aedl}{{\tt channel}!{\tt NOSC\_}} 
in either the initial  flavour or 
the final flavour, like this
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: NOSC\_muon: muon: \#cross: \#energy\\
>}
\end{quote}
%
In this case all diagonal probabilities $P_{\alpha\alpha}$ are unity, and all off-diagonal probablities  $P_{\alpha\beta}$ are zero. This is,
for instance, useful for
neutral current events, since these do not depend on any oscillation 
parameters\footnote{At least in the absence of sterile neutrinos}. The channels
marked as {\tt NOSC\_} are already computed by \GLB{SetRates}
and do not have to be recomputed in the subsequent fit (which calls the undocumented function \GLB{SetNewRates}).
Therefore this feature can be used to speed up the rate 
computation considerably, especially in cases where a large set
of channels exist which are only used for the computation of backgrounds.
Usually it is an excellent approximation to treat backgrounds as if they
were not affected by oscillations\footnote{In the case that the backgrounds have
a sizeable dependence on the oscillation parameters, they carry information
on the oscillation parameters, and therefore behave more like a signal.}.

Note that the energy environment will be described in the next section. 
In addition, one can define pre- and post-smearing effects together
with the channels, which will also be introduced together with the
energy resolution function in the next section.
\index{aedl}{{\tt channel}|)}

@@

\index{aedl}{{\tt energy}|(}
\index{norm}{Energy!resolution|(}
The definition and implementation of the energy resolution function is 
rather sophisticated in \GLOBES . In particular, the choice of the 
proper parameters depends on the experiment and the frequencies of the
 involved neutrino oscillations. This choice also greatly influences the
speed of the calculation.

@@

As illustrated in the figure, an interesting feature in combination with the
channels are pre- and post-smearing effects. Pre-smearing effects are taken into account on the sampling point level, and post-smearing effects on the bin level. Examples for these effects are energy 
dependent efficiencies and (non-beam) backgrounds. Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates. These components can be introduced before or after the integration in \eq~(\ref{eq:simple_int}) is done. 
%
\index{aedl}{{\tt channel}!pre smearing efficiencies@{\tt 
"@pre\_smearing\_efficiencies}}
\index{aedl}{{\tt channel}!pre smearing background@{\tt 
"@pre\_smearing\_background}}
\index{aedl}{{\tt channel}!post smearing efficiencies@{\tt 
"@post\_smearing\_efficiencies}}
\index{aedl}{{\tt channel}!post smearing background@{\tt 
"@post\_smearing\_background}}
%
If they are introduced before, 
we call them
{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
Note that pre-smearing components are always a function of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points. 
Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos. The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}). Examples for post-smearing efficiencies are cuts and detection threshold functions. All post-smearing components have to have as 
many elements as there are energy bins. Efficiencies are multiplicative 
and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
>}
\end{quote}
%
\index{aedl}{{\tt channel}!pre smearing efficiencies@{\tt 
"@pre\_smearing\_efficiencies}}
\index{aedl}{{\tt channel}!pre smearing background@{\tt 
"@pre\_smearing\_background}}
\index{aedl}{{\tt channel}!post smearing efficiencies@{\tt 
"@post\_smearing\_efficiencies}}
\index{aedl}{{\tt channel}!post smearing background@{\tt 
"@post\_smearing\_background}}
%
This experiment uses $10$ sampling points and $5$ bins.

@@

The key idea is to use a ``flat'' model, \ie\ the integrand 
of~\eq~\ref{eq:simple_int}  is well approximated by being piecewise
constant in each sampling step. This is a good approximation as long as
\begin{itemize}
\item
 No details are lost, \ie\ the spacing of sampling points 
is smaller than the energy resolution.
\item
The edges are treated correctly.
\item
 The neutrino oscillations are slow on a scale of the smapling point distance.
\end{itemize}
In this case, \eq~(\ref{eq:simple_int}) is reduced to
\begin{equation}
\label{eq:algo_one}
n_i^c=N/L^2 \, \sum_{j=1}^N \,  \Phi^c(E_j)\,
P^c(E_j)\,
\sigma^c(E_j)\,
K_i^c(E_j) \, \Delta E_j \,.
\end{equation}
The advantages of this algorithm are obvious: All factors
independent of the oscillation parameters have to be only evaluated once
at values of $E$ which are known in advance, which means that they can be put 
into a look-up table. In addition, the probability
has to be only evaluated at previously known values of the energy, which
makes it possible to compute the transition amplitudes for all channels
simultaneously. One assumption is that all involved factors are piece-wise
constant, \ie, they hardly change within each bin. This 
assumption seems to be very restrictive, which is however not quite correct.
First of all, if one analyzes simulated data (which are simulated with the same algorithm), the errors
will cancel between the simulated and fitted data. Second, and more important,
this algorithm is just a very basic integration routine\footnote{It is planed 
for to have something like a Gau\ss-Kronrod scheme as an alternative here.} and converges to
the true result for decreasing step size. Thus if the number of sampling
points is large enough this algorithm is very accurate. 
This algorithm is selected by
%
\index{aedl}{{\tt energy}!type@{\tt "@type}}
%
\begin{quote}
{\tt \tb @type = 1}
\end{quote}
within the {\tt \#energy} environment.
The computation of the bin kernel $K_i^c$ is performed
by \GLOBES. Thus, it requires that the number of bins
{\tt \$bins} and the minimum energy {\tt \$emin} and maximum energy {\tt \$emax} are given in case of equidistant bins.
%
\index{norm}{Energy!resolution}
\index{norm}{Energy!resolution function}
%
As far as the parameterization for the energy resolution function 
$R^c(E,E')$ in \eq~(\ref{eq:kernel}) is concerned, the algorithm uses
a Gau\ss ian
\begin{equation}
R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}} \, .
\end{equation} 
%
\index{aedl}{{\tt energy}!sigma function@{\tt "@sigma\_function}}
%
There are several energy resolution functions available, where by default
{\tt \#standard} is used:
\begin{quote}
{\tt \tb @sigma\_function = \#standard} 
\end{quote}
%
\index{aedl}{{\tt energy}!standard@{\tt \#standard}}
%
The energy resolution function {\tt \#standard} is defined by
\begin{equation}
\label{eq:sigma_e}
\sigma(E)=\alpha\cdot E + \beta \cdot \sqrt{E} +\gamma\, ,
\end{equation}
where the parameters $\alpha, \beta$ and $\gamma$ are provided by the user:
\begin{quote}
{\tt \tb @sigma\_e = \{0.15, 0.0, 0.0\}}
\end{quote}
Currently, another possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta},
%
\index{aedl}{{\tt energy}!inverse beta@{\tt \#inverse\_beta}}
%
which only uses the parameter $\alpha$. It is defined by
\begin{equation}
\sigma(E)= \left\{\begin{array}{cl}
 \alpha \cdot \sqrt{1000}^{-1}\,\sqrt{x-8\cdot10^{-4}}\,,&\mathrm{for}\,\, 
x>1.8\cdot10^{-3}\\
\alpha\cdot10^{-3} \,,&\mathrm{for}\,\, x \leq 1.8\cdot10^{-3}
\end{array} \right.
\end{equation}
The somewhat complicated form is due to the fact that inverse $\beta$-decay
has a neutrino threshold of $1.8\,\mathrm{MeV}$ and that a neutrino
at threshold already produces $\simeq 1\,\mathrm{MeV}$ visible energy in
the detector (for more details see \eg~\cite{Huber:2003pm}).

@@

\index{aedl}{{\tt energy}!type@{\tt  "@type}}
With a type 1 ({\tt @type = 1}) energy resolution 
function, $\sigma_e$ adds on to the energy resolution function 
of the detector $\sigma_c(E)$ such as
\begin{equation}
\sigma_{\mathrm{eff}}(E)^2\simeq \sigma_e^2 + \sigma_c(E)^2\,.
\end{equation} 
Sometimes this behavior is unwanted, and therefore one can try to 
'subtract' the filtering from the energy resolution function by splitting 
the energy resolution function $\sigma(E)_{\mathrm{eff}}$ in
two parts by
\begin{equation}
\sigma_{\mathrm{eff}}(E)^2=\underbrace{\sigma_c(E)^2-\sigma_e^2}_
{\tilde\sigma^2_c(E)}+\sigma_e^2\,,
\end{equation}
where the truncated energy resolution function $\tilde\sigma_c(E)$ 
is used instead of $\sigma_c(E)$ in computing the
smearing data. Thus one obtains as effective energy resolution
\begin{equation}
\sigma_{\mathrm{eff}}(E)^2\simeq \sigma_c(E)^2\,.
\end{equation} 
This scheme is used by choosing as type for the energy resolution
\begin{quote}
{\tt
@type = 2
}
\end{quote}

>>>>

\section{More advanced AEDL features}
\label{sec:advaedl}

@@

\index{aedl}{nuflux@{\tt nuflux}!time@{\tt "@time}}
There are two principal ways to initialize a neutrino flux: 
Either one can use 
a built-in source, or one can provide a file. In both cases,
a flux is defined by the environment {\tt nuflux}, such as
\index{aedl}{nuflux@{\tt nuflux}}
\begin{quote}
  {\tt nuflux(\#name)<\\
\tb $\ldots$\\
\tb @time = 8.0 \\
>}
\end{quote}
with a running time of $8$ years. Note that the running time is used within
the {\tt nuflux} environment. This feature can be used to load the neutrino and antineutrino fluxes separately, in order to combine them with different
running times within one experiment. The name of the flux {\tt \#name} will later be referred to in the channel definitions. Note that \GLOBES\ versions older than 3.0 use the still supported {\tt flux} environment, which is different from {\tt nuflux} by an undocumented normalization factor 5.2 for user-defined fluxes. This difference is explained in the appendix on page~\pageref{app:flux}.

\index{aedl}{nuflux@{\tt nuflux}}

@@

For a built-in neutrino source, one has to specify which
built-in spectrum has to be used, as well as its parameters. The software
will then automatically calculate the neutrino spectrum. Note that in this
case, there is no degree of freedom in the choice of the source units.
The currently available fluxes are described in \Tab~\ref{tab:binfluxes}.
For example, two built-in neutrino factory fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}). In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year. Thus, an example to set up a neutrino factory flux is 
\index{aedl}{nuflux@{\tt nuflux}!builtin@{\tt "@builtin}}
\index{aedl}{nuflux@{\tt nuflux}!parent energy@{\tt "@parent\_energy}}
\index{aedl}{nuflux@{\tt nuflux}!stored muons@{\tt "@stored\_muons}}
\index{aedl}{nuflux@{\tt nuflux}!end point@{\tt "@end\_point}}
\index{aedl}{nuflux@{\tt nuflux}!stored ions@{\tt "@stored\_ions}}
\index{aedl}{nuflux@{\tt nuflux}!gamma@{\tt "@gamma}}
\begin{quote}
{\tt nuflux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}
\end{quote}
MR: SOME MORE ELABORATION AND AN EXAMPLE FOR THE BETA BEAMS MAY BE APPROPRIATE HERE [WW].
%
\index{aedl}{nuflux@{\tt nuflux}!power@{\tt "@power}}
\index{aedl}{nuflux@{\tt nuflux}!norm@{\tt "@norm}}
\index{aedl}{nuflux@{\tt nuflux}!flux file@{\tt "@flux\_file}}
For a user-defined flux, one has to give it the file name:
\begin{quote}
{\tt nuflux(\#user)<}\\
{\tt \tb @flux\_file = "user\_file\_1.dat"\\
\tb @time = 2.0\\
\tb @power = 4.0\\
\tb @norm = 1e+8}\\
{\tt >}
\end{quote}
In this case, the {\tt @norm} variable is an overall normalization which defines a conversion factor from the fluxes in the file to the units in \GLOBES . In general, there are many ways to give the source power of a 
neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
not always straightfoward to calculate. Often, one has to take into account
many things, such as the number of target particles per unit mass. 
In addition, the fluxes will be rescaled by $1/L^2$, which means that the
normalization must contain a factor $L_0^2$. Here $L_0$ is the distance from the source for which the flux is given to the actual neutrino production region. At the end, it is left to the user to ensure that the 
numbers in the flux file give, after the multiplication with {\tt @norm}, 
the proper numbers of produced neutrinos corresponding to the chosen target power {\tt @power}. Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.
For more details on the flux definition, see appendix on page~\pageref{app:flux}.

\index{aedl}{nuflux@{\tt nuflux}!flux file@{\tt "@flux\_file}}\index{norm}{Flux!file}
The software assumes that the given flux file has seven columns and
501 lines with equidistant energies. The format is:
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
In order to access fluxes at arbitrary energies, linear interpolation 
is used by the software. In general, it is advisable to provide the flux between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}), since these values are used by the software. However, if the energy leaves the range of values given in the file, zero is returned.
The the columns for the fluxes for unused flavours have to be filled all the 
same,\eg\ with zeros.

@@

Cross sections\index{norm}{Cross section} will later be used as part of the 
channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
fluxes, they are provided by the user as a data file:
\index{aedl}{cross@{\tt cross}}
\index{aedl}{cross@{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quote}
{\tt cross(\#name)<}\\
{\tt \tb @cross\_file ="user\_file\_1.dat"}\\
{\tt >}
\end{quote}  
This cross section can later be refered to by {\tt \#name}.

Cross sections in \GLOBES\ are given as total cross section divided by energy:
\begin{equation}
\hat\sigma(E)=\sigma(E)/E\,\left[ 10^{-38}\,
\frac{\mathrm{cm}^2}{\mathrm{GeV}} \right]
\end{equation}
The software assumes that the cross section files are text files with 
seven columns and $1001$ lines of the form
\index{norm}{Cross section!file}
\index{aedl}{cross@{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quotation}
$\mathrm{log}_{10} E\quad
\hat\sigma_{\nu_e}\quad
\hat\sigma_{\nu_\mu}\quad
\hat\sigma_{\nu_\tau}\quad
\hat\sigma_{\bar\nu_e}\quad
\hat\sigma_{\bar\nu_\mu}\quad
\hat\sigma_{\bar\nu_\tau}$
\end{quotation}
Here the logarithms of the energy values have to be equidistant. For 
arbitrary energies, linear interpolation is used. If the energy leaves the
range of values given in the file, $0.0$ will be assumed. In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
Unused cross sections have to be filled with zeros, and can not
 be just omitted.

@@

\index{aedl}{channel@{\tt channel}|(}
Channels\index{norm}{Channel} in \GLOBES\ represent an intermediate level 
between the pure oscillation physics given by the oscillation probability
$P_{\alpha\beta}$, and the total event rates composed of signal and 
background. A channel describes the path from one
initial neutrino flavor in the source to the event rates in the detector for one specific interaction type (IT) and final flavor.  
Therefore, a channel contains the description of the 
initial neutrino flavor, its CP eigenvalue 
(neutrino or antineutrino)\footnote{Currently \GLOBES\ does not support lepton
number violating transitions, \ie\ no transitions from neutrino to 
antineutrino (or vice versa) are considered.}, 
the detected neutrino flavor, the interaction cross sections for the chosen interaction type, and the energy resolution function of the detector.

@@

A simple channel definition in \GLOBES\ consists of the flux,
the CP-sign of the initial state, the initial flavor, the final flavor,
the cross sections, and the energy resolution function. In order to refer to
flux, cross sections, and energy resolution functions, they have to be 
defined before with their {\tt \#name} in the respective environments. 
Thus, a simple definition of a channel\index{aedl}{channel@{\tt channel}} is
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}
%
\index{norm}{Oscillation!switching off}
It is also possible to define a channel as no-oscillation by using the
prefix {\tt NOSC\_}\index{aedl}{channel@{\tt channel}!{\tt NOSC\_}} 
in either the initial  flavour or 
the final flavour, like this
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: NOSC\_muon: muon: \#cross: \#energy\\
>}
\end{quote}
%
In this case all diagonal probabilities $P_{\alpha\alpha}$ are unity, and all off-diagonal probablities  $P_{\alpha\beta}$ are zero. This is,
for instance, useful for
neutral current events, since these do not depend on any oscillation 
parameters\footnote{At least in the absence of sterile neutrinos}. The channels
marked as {\tt NOSC\_} are already computed by \GLB{SetRates}
and do not have to be recomputed in the subsequent fit (which calls the undocumented function \GLB{SetNewRates}).
Therefore this feature can be used to speed up the rate 
computation considerably, especially in cases where a large set
of channels exist which are only used for the computation of backgrounds.
Usually it is an excellent approximation to treat backgrounds as if they
were not affected by oscillations\footnote{In the case that the backgrounds have
a sizeable dependence on the oscillation parameters, they carry information
on the oscillation parameters, and therefore behave more like a signal.}.

Note that the energy environment will be described in the next section. 
In addition, one can define pre- and post-smearing effects together
with the channels, which will also be introduced together with the
energy resolution function in the next section.
\index{aedl}{channel@{\tt channel}|)}

@@

\index{aedl}{energy@{\tt energy}|(}
\index{norm}{Energy!resolution|(}
The definition and implementation of the energy resolution function is 
rather sophisticated in \GLOBES . In particular, the choice of the 
proper parameters depends on the experiment and the frequencies of the
 involved neutrino oscillations. This choice also greatly influences the
speed of the calculation.

@@

As illustrated in the figure, an interesting feature in combination with the
channels are pre- and post-smearing effects. Pre-smearing effects are taken into account on the sampling point level, and post-smearing effects on the bin level. Examples for these effects are energy 
dependent efficiencies and (non-beam) backgrounds. Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates. These components can be introduced before or after the integration in \eq~(\ref{eq:simple_int}) is done. 
%
\index{aedl}{channel@{\tt channel}!pre smearing efficiencies@{\tt 
"@pre\_smearing\_efficiencies}}
\index{aedl}{channel@{\tt channel}!pre smearing background@{\tt 
"@pre\_smearing\_background}}
\index{aedl}{channel@{\tt channel}!post smearing efficiencies@{\tt 
"@post\_smearing\_efficiencies}}
\index{aedl}{channel@{\tt channel}!post smearing background@{\tt 
"@post\_smearing\_background}}
%
If they are introduced before, 
we call them
{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
Note that pre-smearing components are always a function of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points. 
Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos. The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}). Examples for post-smearing efficiencies are cuts and detection threshold functions. All post-smearing components have to have as 
many elements as there are energy bins. Efficiencies are multiplicative 
and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
>}
\end{quote}
%
\index{aedl}{channel@{\tt channel}!pre smearing efficiencies@{\tt 
"@pre\_smearing\_efficiencies}}
\index{aedl}{channel@{\tt channel}!pre smearing background@{\tt 
"@pre\_smearing\_background}}
\index{aedl}{channel@{\tt channel}!post smearing efficiencies@{\tt 
"@post\_smearing\_efficiencies}}
\index{aedl}{channel@{\tt channel}!post smearing background@{\tt 
"@post\_smearing\_background}}
%
This experiment uses $10$ sampling points and $5$ bins.

@@

The key idea is to use a ``flat'' model, \ie\ the integrand 
of~\eq~\ref{eq:simple_int}  is well approximated by being piecewise
constant in each sampling step. This is a good approximation as long as
\begin{itemize}
\item
 No details are lost, \ie\ the spacing of sampling points 
is smaller than the energy resolution.
\item
The edges are treated correctly.
\item
 The neutrino oscillations are slow on a scale of the smapling point distance.
\end{itemize}
In this case, \eq~(\ref{eq:simple_int}) is reduced to
\begin{equation}
\label{eq:algo_one}
n_i^c=N/L^2 \, \sum_{j=1}^N \,  \Phi^c(E_j)\,
P^c(E_j)\,
\sigma^c(E_j)\,
K_i^c(E_j) \, \Delta E_j \,.
\end{equation}
The advantages of this algorithm are obvious: All factors
independent of the oscillation parameters have to be only evaluated once
at values of $E$ which are known in advance, which means that they can be put 
into a look-up table. In addition, the probability
has to be only evaluated at previously known values of the energy, which
makes it possible to compute the transition amplitudes for all channels
simultaneously. One assumption is that all involved factors are piece-wise
constant, \ie, they hardly change within each bin. This 
assumption seems to be very restrictive, which is however not quite correct.
First of all, if one analyzes simulated data (which are simulated with the same algorithm), the errors
will cancel between the simulated and fitted data. Second, and more important,
this algorithm is just a very basic integration routine\footnote{It is planed 
for to have something like a Gau\ss-Kronrod scheme as an alternative here.} and converges to
the true result for decreasing step size. Thus if the number of sampling
points is large enough this algorithm is very accurate. 
This algorithm is selected by
%
\index{aedl}{energy@{\tt energy}!type@{\tt "@type}}
%
\begin{quote}
{\tt \tb @type = 1}
\end{quote}
within the {\tt \#energy} environment.
The computation of the bin kernel $K_i^c$ is performed
by \GLOBES. Thus, it requires that the number of bins
{\tt \$bins} and the minimum energy {\tt \$emin} and maximum energy {\tt \$emax} are given in case of equidistant bins.
%
\index{norm}{Energy!resolution}
\index{norm}{Energy!resolution function}
%
As far as the parameterization for the energy resolution function 
$R^c(E,E')$ in \eq~(\ref{eq:kernel}) is concerned, the algorithm uses
a Gau\ss ian
\begin{equation}
R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}} \, .
\end{equation} 
%
\index{aedl}{energy@{\tt energy}!sigma function@{\tt "@sigma\_function}}
%
There are several energy resolution functions available, where by default
{\tt \#standard} is used:
\begin{quote}
{\tt \tb @sigma\_function = \#standard} 
\end{quote}
%
\index{aedl}{energy@{\tt energy}!standard@{\tt \#standard}}
%
The energy resolution function {\tt \#standard} is defined by
\begin{equation}
\label{eq:sigma_e}
\sigma(E)=\alpha\cdot E + \beta \cdot \sqrt{E} +\gamma\, ,
\end{equation}
where the parameters $\alpha, \beta$ and $\gamma$ are provided by the user:
\begin{quote}
{\tt \tb @sigma\_e = \{0.15, 0.0, 0.0\}}
\end{quote}
Currently, another possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta},
%
\index{aedl}{energy@{\tt energy}!inverse beta@{\tt \#inverse\_beta}}
%
which only uses the parameter $\alpha$. It is defined by
\begin{equation}
\sigma(E)= \left\{\begin{array}{cl}
 \alpha \cdot \sqrt{1000}^{-1}\,\sqrt{x-8\cdot10^{-4}}\,,&\mathrm{for}\,\, 
x>1.8\cdot10^{-3}\\
\alpha\cdot10^{-3} \,,&\mathrm{for}\,\, x \leq 1.8\cdot10^{-3}
\end{array} \right.
\end{equation}
The somewhat complicated form is due to the fact that inverse $\beta$-decay
has a neutrino threshold of $1.8\,\mathrm{MeV}$ and that a neutrino
at threshold already produces $\simeq 1\,\mathrm{MeV}$ visible energy in
the detector (for more details see \eg~\cite{Huber:2003pm}).

@@

\index{aedl}{energy@{\tt energy}!type@{\tt  "@type}}
With a type 1 ({\tt @type = 1}) energy resolution 
function, $\sigma_e$ adds on to the energy resolution function 
of the detector $\sigma_c(E)$ such as
\begin{equation}
\sigma_{\mathrm{eff}}(E)^2\simeq \sigma_e^2 + \sigma_c(E)^2\,.
\end{equation} 
Sometimes this behavior is unwanted, and therefore one can try to 
'subtract' the filtering from the energy resolution function by splitting 
the energy resolution function $\sigma(E)_{\mathrm{eff}}$ in
two parts by
\begin{equation}
\sigma_{\mathrm{eff}}(E)^2=\underbrace{\sigma_c(E)^2-\sigma_e^2}_
{\tilde\sigma^2_c(E)}+\sigma_e^2\,,
\end{equation}
where the truncated energy resolution function $\tilde\sigma_c(E)$ 
is used instead of $\sigma_c(E)$ in computing the
smearing data. Thus one obtains as effective energy resolution
\begin{equation}
\sigma_{\mathrm{eff}}(E)^2\simeq \sigma_c(E)^2\,.
\end{equation} 
This scheme is used by choosing as type for the energy resolution
\begin{quote}
{\tt
@type = 2
}
\end{quote}
